Flatiron cleaning

### 2021-12-01 Rmd/rdi analysis.Rmd
reorganize code clean up, complete step 2 and start step 3

```{r drugs episode cleanup}
drugs <- drugs %>% 
  filter(episodedatasource == "Administrations" #& 
           # ismaintenancetherapy == "FALSE" # keep maintenance for a correct recoding of line number
         ) %>%
  # bind with clinical
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate)
```

@@

- 30 days rule excluding patient who didn’t survive for at least 30 days after diagnosis  
This is something that Flatiron strongly recommends, but does not require. Flatiron proposal reviewers will likely provide comments on this if not incorporated into a proposal.

@@

Are excluded :  
-	Borderline cases (not invasive ovarian cancer)  
-	Data from one PracticeID  (Flatiron requested removal of this practice ID from the dataset)  
-	Certain historical structured data were not available for patients seen at academic sites prior to 2015 – Flatiron has asked that if conducting survival analyses, these patients are removed (patients diagnosed < 2015 at academic sites)

```{r rule between diagnosis and treatment}
drugs1 <- drugs %>% 
  # 1. 60 days rule between diagnosis and treatment # Loss 1
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "months")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "months")) %>% 
  ungroup() %>% 
  mutate(thirty_days_exclusion = interval(start = diagnosisdate, end = dateofdeath)/
           duration(n=1, units = "days")
  )
```

```{r}
drugs1 <- drugs1 %>% 
  mutate(flatiron_rules = case_when(
    days_at_first_drug > 90 | 
      days_at_surg > 90                          ~ "exclude",
    thirty_days_exclusion <= 30                  ~ "exclude",
    exclude == "1"                               ~ "exclude"
  )) %>%
  group_by(patientid) %>% 
  fill(flatiron_rules, .direction = "downup")

@@

## 2. step 2: Filter patients with cinical characteristics fitting to the study

- exclude patients whose histology is unknown  
- remove chemotherapy only patients
```{r}
drugs1 <- drugs1 %>% 
  filter(!is.na(histology)) # loss 287 ################## Will do when bind to drugs
```

@@

@@

- exclude the patients who had surgery but we don't have a date of surgery

```{r}
drugs1 <- drugs1 %>% 
  filter(!(issurgery == "Yes" & is.na(surgerydate))) 
```

@@

#### Combine multiple administrations of the same drug at the same date
```{r combine amount same drug one day}
drugs1 <- drugs1 %>% 
  #When want to remove patients who didn't have taxel/platin as 1st line -> remove 260 patients
  # filter(!(linenumber == 1 & !str_detect(linename, "taxel|platin"))) %>% 
  # filter(str_detect(drugname, "taxel|platin") & str_detect(route, "venous|peritoneal")) %>% ############## LATER------------!!!!!
  select(-c(enhancedcohort, episodedatasource, drugcategory, detaileddrugcategory)) %>%

  # 3. Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
# There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.
# For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.
# Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend. 
# Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate, 
           drugname, units, surgerydate, days_at_first_drug, days_at_surg, 
           diagnosisdate, ismaintenancetherapy) %>%
  summarise_at(vars(route, amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  mutate(route = case_when(
    str_detect(route, ";")            ~ "Intravenous",
    TRUE                              ~ route
  )) %>% 
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  ungroup() %>% 
  mutate(across(starts_with("amount"), ~ na_if(., "NA"))) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  mutate(across(starts_with("amount"), ~ as.numeric(.))) %>% 
  # mutate(amount_1 = as.numeric(amount_1), amount_2 = as.numeric(amount_2),
  #        amount_3 = as.numeric(amount_3), amount_4 = as.numeric(amount_4)) %>%
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

#### Recode linenumber depending on surgery date (Categorize Neo and Adjuavnt)
```{r line coding}
therapy_line <- drugs1 %>% 
  # 4. Define Adjuvant, Neoadjuvant, chemo or surgery----
  mutate(chemotherapy_type = case_when(
    # ismaintenancetherapy == "TRUE"    ~ "maintenance",
    episodedate <= surgerydate        ~ "Neoadjuvant", # F63ABD6AEB12C, F4F7B1C5DF24F intraperitoneal Carboplatin
    episodedate > surgerydate         ~ "Adjuvant",
    is.na(surgerydate) &
      !is.na(episodedate)             ~ "Chemotherapy only",
    !is.na(surgerydate) &
      is.na(episodedate)              ~ "Surgery only"
    )) %>% 
  mutate(chemotherapy_type = 
           factor(chemotherapy_type, levels = 
                    c("Neoadjuvant", "Adjuvant", "Chemotherapy only", "Surgery only"))) %>%
  group_by(patientid) %>% 
  mutate(had_neo = ifelse(str_detect(chemotherapy_type, "Neo"), "Yes", NA_character_)) %>% 
  fill(had_neo, .direction = "updown") %>% 
  mutate(had_adj = ifelse(str_detect(chemotherapy_type, "Adj"), "Yes", NA_character_)) %>% 
  fill(had_adj, .direction = "updown") %>% 
  mutate(treatment_sequence = case_when(
    had_neo == "Yes" &
      had_adj == "Yes"                           ~ "neo + surg + adj",
    had_neo == "Yes" &
      is.na(had_adj)                             ~ "neo + surg",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "surg + adj",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only"
  )) %>% 
  mutate(therapy = case_when(
    chemotherapy_type == "Neoadjuvant"           ~ "Upfront Chemo",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "Upfront Surgery"
  )) %>% 
  fill(therapy, .direction = "downup") %>% 
  # mutate(therapy1 = case_when(
  #   chemotherapy_type == "Adjuvant"              ~ "Upfront Surgery",
  # )) %>% 
  # fill(therapy1, .direction = "downup") %>% 
  # mutate(therapy = coalesce(therapy, therapy1)) %>% 
  # select( -therapy1) %>%

  # 7.Calculate cycle for each line----
  mutate(cycle_drugname = str_remove(drugname, " protein-bound")) %>%  # F003F2BEEDB37 need to change drugname pacli = pacli bp
  group_by(patientid, linenumber_new, cycle_drugname) %>% # Wrong line name F001443D8B85C ~~~~~~~~~~~~~~~~~~~~~~~~~~ Fix later QUESTION Dr Chern is ok
  mutate(drugname_count_perline = n()) %>%  # F0000AD999ABF
  ungroup() %>% 
  mutate(across(c("linenumber", "linenumber_new", "linenumber_adj"), as.factor)) %>% 
  group_by(patientid, linenumber_new) %>% 
  mutate(linestartdate_new = min(episodedate)) %>% 
  mutate(lineenddate_new = max(episodedate)) %>% 
  ungroup() %>% 
  filter(treatment_sequence == "neo + surg + adj" |
           treatment_sequence == "surg + adj")

@@

### step 6: remove patients with long gap between surgery and adjuvant start ????
Depending on the answers:

#### remove patients with more than 9 cycles total in upfront chemo or upfront surgery
Did a gap of 3 months
```{r}
remove_delayed_drug <- therapy_line %>%
  mutate(interval = case_when(
      chemotherapy_type == "Adjuvant" &
        linenumber_adj == 1       
      ~ interval(start = surgerydate, end = linestartdate_new) / duration(n = 1, units = "months")
    ),
  ) %>% 
  arrange(patientid, episodedate) %>%
  filter(interval > 3) %>%
  distinct(patientid, .keep_all = TRUE)

therapy_line <- therapy_line %>%
  # Remove patients with more than 1 line as Neo
  filter(!str_detect(patientid, paste(remove_delayed_drug$patientid, collapse = "|")))
therapy_line %>% distinct(patientid) %>% nrow()
```

@@

#### Code cycle
```{r cycle coding}
# Code cycle number
therapy_cycle <- therapy_line %>% 
  filter(str_detect(drugname, "taxel|platin")) %>% # Need to that now for patient who got Bev only once per line F02D8628A5394

  arrange(patientid, episodedate) %>% 
  group_by(patientid, linenumber_new) %>%
  # Need the number of time of the "major" drug is seen in the line
  mutate(cycle_count_perline = min(drugname_count_perline)) %>%
  mutate(cycle_major_drugname = case_when(
    cycle_count_perline == drugname_count_perline          ~ cycle_drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  group_by(patientid, linenumber_new, cycle_major_drugname) %>%
  mutate(cycle_increment = row_number(cycle_major_drugname)) %>% # Look at F002DD2953889 had a long time without one drug------------------
  # F01D908A013B7 has more pacli at the end but in the same line

  # Look at dose dense
  group_by(patientid, linenumber_new, cycle_increment, cycle_drugname) %>%  # weird dose dense F064D9EB733B3 F050B815FC126
  mutate(dose_dense = n()) %>% 
  mutate(dose_dense = max(dose_dense)) %>%  # too much cycle
  ungroup() # %>% 
  # select(patientid, linename, linenumber_new, drugname, episodedate, chemotherapy_type, cycle_increment, dose_dense, amount, ismaintenancetherapy,
  #        is_dose_dense,cycle_major_drugname, cycle_drugname,
  #        cycle_count_perline, cycle_major_drugname, drugname_count_perline, cycle_start_date) #%>%  # F002DD2953889, F003F2BEEDB37 only 1+2

@@

### step 10: Bind everything together and calculating target_auc and expected dose

(hold on to the thought that what I call `target_auc` include the target AUC calculated for Carboplatin and target dose calculated for cisplatin and taxanes, they are just present in the same column called `target_auc` for a tidy data organisation - You can find the calculations formula at the end of the document)  
When I bind the data together (drug + corresponding creatinine + ...), I take the closest date of creatinine, height, weight, BSA, AUC, from the cycle date to calculate `target_auc` for Carboplatin, cisplatin and taxanes and their `expected_dose`. It can be days in between but it can also be more.

>>>>

- Select episodedatasource == "Administrations"
```{r drugs episode cleanup}
drugs <- drugs %>% 
  filter(episodedatasource == "Administrations" #& 
           # ismaintenancetherapy == "FALSE" # keep maintenance for a correct recoding of line number
         ) %>%
  # bind with clinical
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate)
```

- Combine multiple administrations of the same drug at the same date
Flatiron says:
There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.  
For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.  
Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend.   
Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
```{r combine amount same drug one day}
drugs1 <- drugs %>% 
  # 3. Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
# There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.
# For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.
# Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend. 
# Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate, 
           drugname, units, ismaintenancetherapy) %>%
  summarise_at(vars(route, amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  mutate(route = case_when(
    str_detect(route, ";")            ~ "Intravenous",
    TRUE                              ~ route
  )) %>% 
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  ungroup() %>% 
  mutate(across(starts_with("amount"), ~ na_if(., "NA"))) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  mutate(across(starts_with("amount"), ~ as.numeric(.))) %>% 
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 so remove
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

- Recode linenumber depending on surgery date (Categorize Neo and Adjuavnt)
```{r line coding}
therapy_line <- drugs1 %>% 
  # bind again with clinical - easier that way
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate) %>% 
  # 4. Define Adjuvant, Neoadjuvant, chemo or surgery----
  mutate(chemotherapy_type = case_when(
    # ismaintenancetherapy == "TRUE"    ~ "maintenance",
    episodedate <= surgerydate        ~ "Neoadjuvant", # F63ABD6AEB12C, F4F7B1C5DF24F intraperitoneal Carboplatin
    episodedate > surgerydate         ~ "Adjuvant",
    is.na(surgerydate) &
      !is.na(episodedate)             ~ "Chemotherapy only",
    !is.na(surgerydate) &
      is.na(episodedate)              ~ "Surgery only"
    )) %>% 
  mutate(chemotherapy_type = 
           factor(chemotherapy_type, levels = 
                    c("Neoadjuvant", "Adjuvant", "Chemotherapy only", "Surgery only"))) %>%
  group_by(patientid) %>% 
  mutate(had_neo = ifelse(str_detect(chemotherapy_type, "Neo"), "Yes", NA_character_)) %>% 
  fill(had_neo, .direction = "updown") %>% 
  mutate(had_adj = ifelse(str_detect(chemotherapy_type, "Adj"), "Yes", NA_character_)) %>% 
  fill(had_adj, .direction = "updown") %>% 
  mutate(treatment_sequence = case_when(
    had_neo == "Yes" &
      had_adj == "Yes"                           ~ "neo + surg + adj",
    had_neo == "Yes" &
      is.na(had_adj)                             ~ "neo + surg",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "surg + adj",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only"
  )) %>% 
  mutate(therapy = case_when(
    chemotherapy_type == "Neoadjuvant"           ~ "Upfront Chemo",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "Upfront Surgery"
  )) %>% 
  fill(therapy, .direction = "downup") %>%

  # 7.Calculate cycle for each line----
  mutate(cycle_drugname = str_remove(drugname, " protein-bound")) %>%  # F003F2BEEDB37 need to change drugname pacli = pacli bp
  group_by(patientid, linenumber_new, cycle_drugname) %>% # Wrong line name F001443D8B85C ~~~~~~~~~~~~~~~~~~~~~~~~~~ Fix later QUESTION Dr Chern is ok
  mutate(#drugname_count_perline = n(),
         drugname_count_perline = case_when(
           str_detect(drugname, "platin|taxel")     ~ n()
         )
           ) %>%  # F0000AD999ABF
  ungroup() %>% 
  mutate(across(c("linenumber", "linenumber_new", "linenumber_adj"), as.factor)) %>% 
  group_by(patientid, linenumber_new) %>% 
  mutate(linestartdate_new = min(episodedate)) %>% 
  mutate(lineenddate_new = max(episodedate)) %>% 
  ungroup()
```

- Code cycle number
```{r cycle coding}
# Code cycle number
therapy_cycle <- therapy_line %>% 
  # filter(str_detect(drugname, "taxel|platin")) %>% # Need to that now for patient who got Bev only once per line F02D8628A5394

  arrange(patientid, episodedate) %>% 
  group_by(patientid, linenumber_new) %>%
  # Need the number of time of the "major" drug is seen in the line
  mutate(cycle_count_perline = min(drugname_count_perline, na.rm = TRUE)) %>%
  mutate(cycle_major_drugname = case_when(
    cycle_count_perline == drugname_count_perline          ~ cycle_drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  group_by(patientid, linenumber_new, cycle_major_drugname) %>%
  mutate(cycle_increment = row_number(cycle_major_drugname)) %>% # Look at F002DD2953889 had a long time without one drug------------------
  # F01D908A013B7 has more pacli at the end but in the same line

  # Look at dose dense
  group_by(patientid, linenumber_new, cycle_increment, cycle_drugname) %>%  # weird dose dense F064D9EB733B3 F050B815FC126
  mutate(dose_dense = n()) %>% 
  mutate(dose_dense = max(dose_dense)) %>%  # too much cycle
  ungroup()

# rm(drugs, drugs1, therapy_line)
```

- Bind everything together to calculate target auc/target dose and expected dose

(hold on to the thought that what I call `target_auc` include the target AUC calculated for Carboplatin and target dose calculated for cisplatin and paclitaxel, they are just present in the same column called `target_auc` for a tidy data organisation - You can find the calculations formula at the end of the document)  
When I bind the data together (drug + corresponding creatinine + ...), I take the closest date of creatinine, height, weight, BSA, AUC, from the cycle date to calculate `target_auc` for Carboplatin, and paclitaxel and their `expected_dose`. It can be days in between but it can also be more.

  mutate(actual_bmi = weight / (height * height)) %>%
  mutate(actual_bmi_cat = case_when(
    actual_bmi < 25                    ~ "Underweight and normal weight",
    actual_bmi >= 25 &
      actual_bmi < 30                  ~ "Overweight",
    actual_bmi >= 30                   ~ "Obese"
  )) %>%
  mutate(actual_bmi_cat = factor(actual_bmi_cat, 
                                  levels = c("Underweight and normal weight", "Overweight", "Obese"))) %>%
  # select(patientid, weight, height, bmi_at_dx, ideal_weight, adjusted_weight, adjusted_bmi)
  # Calculate CrCl
  # inner_join(., clinical_data %>% select(-surgerydate),
  #            by = "patientid") %>% 
  mutate(CrCl = (
       (
    ((140 - ageatdx) * weight) /
    (72 * creatinine)
       ) *
    0.85)
  ) %>% 
  mutate(CrCl = case_when(
    CrCl > 125                        ~ 125,
    TRUE                              ~ CrCl
  )) %>% 
  mutate(adjusted_CrCl = case_when(
    actual_bmi >= 25                  ~ CrCl,
    actual_bmi < 25                   ~ (((140 - ageatdx) * adjusted_weight) / (72 * creatinine)) * 0.85
  )) %>% 
  mutate(adjusted_CrCl = case_when(
    adjusted_CrCl > 125               ~ 125,
    TRUE                              ~ adjusted_CrCl
  )) %>%

# Variable for filtering

```{r rule between diagnosis and treatment}
Treatment <- Treatment %>% 
  # 1. 60 days rule between diagnosis and treatment # Loss 1
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "days")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "days")) %>% 
  ungroup() %>% 
  mutate(thirty_days_exclusion = interval(start = diagnosisdate, end = dateofdeath)/
           duration(n=1, units = "days")
  ) %>% 
  group_by(patientid) %>% 
  mutate(had_carbo_pacli = case_when(
    drugname == "paclitaxel" |
      drugname == "carboplatin"             ~ "Yes"
  )) %>% 
  fill(had_carbo_pacli, .direction = "updown") %>% 
  ungroup() %>% 
  mutate(interval_surgery_to_adj_start = case_when(
      chemotherapy_type == "Adjuvant" &
        linenumber_adj == 1       
      ~ interval(start = surgerydate, end = linestartdate_new) / duration(n = 1, units = "months")
    ),
  )
```

```{r RDI calculation}
Frontline <- Treatment %>% 
  # filter(chemotherapy_type == "Neoadjuvant" |
  #          (chemotherapy_type == "Adjuvant" &
  #             linenumber_adj == 1)) %>%

  # Average RDI by drug by line (will NOT combine pacli-doce and cis-carbo-oxalo)
  group_by(patientid, chemotherapy_type, drugname) %>%
  # mutate(mean_RDI_per_drug = case_when(
  #   chemotherapy_type == "Adjuvant" &
  #     linenumber_adj == 1                  ~ round(mean(relative_dose_intensity, na.rm = TRUE), 3),
  #   chemotherapy_type == "Neoadjuvant"     ~ round(mean(relative_dose_intensity, na.rm = TRUE), 3)
  # )) %>% 
  mutate(mean_RDI_per_drug = round(mean(relative_dose_intensity, na.rm = TRUE), 3)
  ) %>% 
  ungroup() %>% 
  distinct(patientid, chemotherapy_type, drugname, .keep_all = TRUE)
```

```{r RDI calculation group}
Frontline <- Frontline %>% 
  # general
  mutate(RDI_grp = case_when(
    mean_RDI_per_drug >= 0.85              ~ "Adequate dosing",
    mean_RDI_per_drug < 0.85               ~ "Underdosed"
  )) %>% 
  mutate(RDI_grp = factor(RDI_grp, 
                           levels = c("Adequate dosing", "Underdosed"))) %>% 
  mutate(RDI_grp1 = case_when(
    mean_RDI_per_drug < 0.75               ~ "Highly underdosed",
    mean_RDI_per_drug >= 0.75 &
      mean_RDI_per_drug < 0.85             ~ "Underdosed",
    mean_RDI_per_drug >= 0.85 &
      mean_RDI_per_drug < 1.1              ~ "Adequate dosing",
    mean_RDI_per_drug >= 1.1               ~ "Overdosing"
  )) %>% 
  mutate(RDI_grp1 = factor(RDI_grp1, 
                          levels = c("Adequate dosing",
                                     "Underdosed", 
                                     "Highly underdosed",
                                     "Overdosing"))) %>% 
  # separated
  mutate(RDI_carbo_neo = case_when(
    drugname == "carboplatin" &
      chemotherapy_type == "Neoadjuvant"    ~ mean_RDI_per_drug,
    TRUE                                    ~ NA_real_
  )) %>% 
  mutate(RDI_pacli_neo = case_when(
    drugname == "paclitaxel" &
      chemotherapy_type == "Neoadjuvant"    ~ mean_RDI_per_drug,
    TRUE                                    ~ NA_real_
  )) %>% 
  mutate(RDI_carbo_adj = case_when(
    drugname == "carboplatin" &
      chemotherapy_type == "Adjuvant"    ~ mean_RDI_per_drug,
    TRUE                                    ~ NA_real_
  )) %>% 
  mutate(RDI_pacli_adj = case_when(
    drugname == "paclitaxel" &
      chemotherapy_type == "Adjuvant"    ~ mean_RDI_per_drug,
    TRUE                                    ~ NA_real_
  )) %>%

  mutate(RDI_carbo_neo_grp = case_when(
    RDI_carbo_neo >= 0.85              ~ "Adequate dosing",
    RDI_carbo_neo < 0.85               ~ "Underdosed"
  )) %>% 
  mutate(RDI_carbo_neo_grp = factor(RDI_carbo_neo_grp, 
                           levels = c("Adequate dosing", "Underdosed"))) %>% 
  mutate(RDI_carbo_neo_grp1 = case_when(
    RDI_carbo_neo < 0.75               ~ "Highly underdosed",
    RDI_carbo_neo >= 0.75 &
      RDI_carbo_neo < 0.85             ~ "Underdosed",
    RDI_carbo_neo >= 0.85 &
      RDI_carbo_neo < 1.1              ~ "Adequate dosing",
    RDI_carbo_neo >= 1.1               ~ "Overdosing"
  )) %>% 
  mutate(RDI_carbo_neo_grp1 = factor(RDI_carbo_neo_grp1, 
                          levels = c("Adequate dosing",
                                     "Underdosed", 
                                     "Highly underdosed",
                                     "Overdosing"))) %>% 
  mutate(RDI_pacli_neo_grp = case_when(
    RDI_pacli_neo >= 0.85              ~ "Adequate dosing",
    RDI_pacli_neo < 0.85               ~ "Underdosed"
  )) %>% 
  mutate(RDI_pacli_neo_grp = factor(RDI_pacli_neo_grp, 
                           levels = c("Adequate dosing", "Underdosed"))) %>% 
  mutate(RDI_pacli_neo_grp1 = case_when(
    RDI_pacli_neo < 0.75               ~ "Highly underdosed",
    RDI_pacli_neo >= 0.75 &
      RDI_pacli_neo < 0.85             ~ "Underdosed",
    RDI_pacli_neo >= 0.85 &
      RDI_pacli_neo < 1.1              ~ "Adequate dosing",
    RDI_pacli_neo >= 1.1               ~ "Overdosing"
  )) %>% 
  mutate(RDI_pacli_neo_grp1 = factor(RDI_pacli_neo_grp1, 
                          levels = c("Adequate dosing",
                                     "Underdosed", 
                                     "Highly underdosed",
                                     "Overdosing"))) %>%

  mutate(RDI_carbo_adj_grp = case_when(
    RDI_carbo_adj >= 0.85              ~ "Adequate dosing",
    RDI_carbo_adj < 0.85               ~ "Underdosed"
  )) %>% 
  mutate(RDI_carbo_adj_grp = factor(RDI_carbo_adj_grp, 
                           levels = c("Adequate dosing", "Underdosed"))) %>% 
  mutate(RDI_carbo_adj_grp1 = case_when(
    RDI_carbo_adj < 0.75               ~ "Highly underdosed",
    RDI_carbo_adj >= 0.75 &
      RDI_carbo_adj < 0.85             ~ "Underdosed",
    RDI_carbo_adj >= 0.85 &
      RDI_carbo_adj < 1.1              ~ "Adequate dosing",
    RDI_carbo_adj >= 1.1               ~ "Overdosing"
  )) %>% 
  mutate(RDI_carbo_adj_grp1 = factor(RDI_carbo_adj_grp1, 
                          levels = c("Adequate dosing",
                                     "Underdosed", 
                                     "Highly underdosed",
                                     "Overdosing"))) %>% 
  mutate(RDI_pacli_adj_grp = case_when(
    RDI_pacli_adj >= 0.85              ~ "Adequate dosing",
    RDI_pacli_adj < 0.85               ~ "Underdosed"
  )) %>% 
  mutate(RDI_pacli_adj_grp = factor(RDI_pacli_adj_grp, 
                           levels = c("Adequate dosing", "Underdosed"))) %>% 
  mutate(RDI_pacli_adj_grp1 = case_when(
    RDI_pacli_adj < 0.75               ~ "Highly underdosed",
    RDI_pacli_adj >= 0.75 &
      RDI_pacli_adj < 0.85             ~ "Underdosed",
    RDI_pacli_adj >= 0.85 &
      RDI_pacli_adj < 1.1              ~ "Adequate dosing",
    RDI_pacli_adj >= 1.1               ~ "Overdosing"
  )) %>% 
  mutate(RDI_pacli_adj_grp1 = factor(RDI_pacli_adj_grp1, 
                          levels = c("Adequate dosing",
                                     "Underdosed", 
                                     "Highly underdosed",
                                     "Overdosing")))
```

```{r Frontline_adjusted}
Frontline_adjusted <- Frontline %>% 
  group_by(patientid) %>% 
  fill(RDI_carbo_neo, RDI_carbo_neo_grp, RDI_carbo_neo_grp1, 
       RDI_pacli_neo, RDI_pacli_neo_grp, RDI_pacli_neo_grp1, 
       RDI_carbo_adj, RDI_carbo_adj_grp, RDI_carbo_adj_grp1, 
       RDI_pacli_adj, RDI_pacli_adj_grp, RDI_pacli_adj_grp1, 
       .direction = "downup") %>% 
  distinct(patientid, .keep_all = TRUE)
```

```{r}
rm(areaUC, body_surface_area, creatinine, clinical_data, drugs, drugs1, 
   height#, therapy_cycle, therapy_line, Treatment, weight
   )
```

# **Number of patients : `r Frontline_adjusted %>% distinct(patientid) %>% nrow()`**

######################################################################################################################
######################################################################################################################
######################################################################################################################
######################################################################################################################

<!-- # Variable for filtering -->

<!-- ```{r rule between diagnosis and treatment} -->
<!-- drugs1 <- Frontline_adjusted %>%  -->
<!--   # 1. 60 days rule between diagnosis and treatment # Loss 1 -->
<!--   group_by(patientid) %>%  -->
<!--   mutate(first_drug_date = first(episodedate)) %>%  -->
<!--   mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/ -->
<!--            duration(n=1, units = "days")) %>%  -->
<!--   mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/ -->
<!--            duration(n=1, units = "days")) %>%  -->
<!--   ungroup() %>%  -->
<!--   mutate(thirty_days_exclusion = interval(start = diagnosisdate, end = dateofdeath)/ -->
<!--            duration(n=1, units = "days") -->
<!--   ) -->
<!-- ``` -->

@@

- 30 days rule excluding patient who didn't survive for at least 30 days after diagnosis  
This is something that Flatiron strongly recommends, but does not require. Flatiron proposal reviewers will likely provide comments on this if not incorporated into a proposal.

@@

Are excluded :  
-	Borderline cases (not invasive ovarian cancer)  
-	Data from one PracticeID  (Flatiron requested removal of this practice ID from the dataset)  
-	Certain historical structured data were not available for patients seen at academic sites prior to 2015 – Flatiron has asked that if conducting survival analyses, these patients are removed (patients diagnosed < 2015 at academic sites)  
```{r}
drugs1 <- Frontline_adjusted %>% 
  mutate(flatiron_rules = case_when(
    days_at_first_drug > 90 | 
      days_at_surg > 90                          ~ "exclude",
    thirty_days_exclusion <= 30                  ~ "exclude",
    exclude == "1"                               ~ "exclude"
  )) %>%
  group_by(patientid) %>% 
  fill(flatiron_rules, .direction = "downup")

@@

## 2. step 2: Filter patients with cinical characteristics fitting to the study criteria

- exclude patients whose histology is unknown  
- remove chemotherapy only patients
- include only patients who received platin and/or taxane
```{r}
drugs1 <- drugs1 %>% 
  filter(!is.na(histology)) # loss 287 ################## Will do when bind to drugs
```

@@

```{r}
drugs1 <- drugs1 %>%
  filter(had_carbo_pacli == "Yes")
```

@@

- exclude the patients who had surgery but we don't have a date of surgery -> NONE anymore

```{r}
drugs11 <- drugs1 %>% 
  filter(!(issurgery == "Yes" & is.na(surgerydate))) 
```

- remove patients with long gap between surgery and adjuvant start -> NONE anymore  
Did a 3 months gap
```{r}
drugs11 <- drugs11 %>%
  arrange(patientid, episodedate) %>%
  mutate(ex = case_when(
    interval_surgery_to_adj_start > 3                  ~ "exclude"
  )) %>% 
  filter(is.na(ex))

drugs11 %>% distinct(patientid) %>% nrow()
```

RESTART HERE ----------------------------------------------------

@@

```{r}
therapy_line <- therapy_line %>% 
  filter(treatment_sequence == "neo + surg + adj" |
           treatment_sequence == "surg + adj")

@@

#### remove patients with more than 9 cycles total in upfront chemo or upfront surgery

@@

@@
